---
title: "Chapter_2_Exercises"
format: html
editor: visual
embed-resources: true
---

## Answers to Chapter Two exercises.

### See also https://xcelab.net/rmpubs/rethinking/rethinking_solutions_2.pdf

2E1. Which of the expressions below correspond to the statement: The probability of rain on Monday?

```{r}
# Answer: Pr(rain|Monday)
```

2E2. Which of the following statements corresponds to the expression: Pr(Monday\|rain)?

```{r}
# Answer: 3) The probability that it is Monday, given that it is raining
```

2E3. Which of the expressions below correspond to the statement: The probability that it is Monday, given that it is raining?

```{r}
# Answer: 1) Pr(Monday|rain)

```

2E4.

```{r}
# Answer: There is a true value of the amount of water on earth # and for each individual spot on earth it will be either covered in water, or it won't be. However, if we take a random sample, it will more than likely be water. 
```

2M1. Compute and plot the grid approximate posterior distribution for each of the following set of observations. Assume a uniform prior.

```{r}
# Answer. Let's make a function that will do this for us 
# Arguments for the function? n = tosses, w = number of waters 
grid_post <- function(n , w){
  # define grid
  p_grid <- seq(from = 0, to = 1, length.out = 100)
  
  # define uniform prior
  prior <- rep(1, 100)
  
  # compute likliehood at each value in the grid. We put in the arguments w and n here to calculate the probability w waters in n tosses with a p of ps. 
  likelihood <- dbinom(w, n, prob = p_grid)
 
  # compute prior x liklihood
  unstd.posterior <- likelihood*prior
  
  # standardize so it sums to 1 
  posterior <- unstd.posterior / sum(unstd.posterior)
  
  plot(p_grid, posterior, type = "b", xlab = "probability of water",
       ylab = "posterior probability")
}
grid_post(w = 3, n = 3)
grid_post(w = 3, n = 4)
grid_post(w = 5, n = 7)
```

2M2. Do the same but assume a prior = 0 when p \< 0.5 and is a positive constant when p \> 0.5

```{r}
# Answer. Let's make a function that will do this for us 
# Arguments for the function? n = tosses, w = number of waters 
grid_post_step_prior <- function(n , w){
  # define grid
  p_grid <- seq(from = 0, to = 1, length.out = 100)
  
  # define prior 
  prior <- ifelse(p_grid < 0.5, 0, 1)
  
  # compute likliehood at each value in the grid. We put in the arguments w and n here to calculate the probability w waters in n tosses with a p of ps. 
  likelihood <- dbinom(w, n, prob = p_grid)
 
  # compute prior x liklihood
  unstd.posterior <- likelihood*prior
  
  # standardize so it sums to 1 
  posterior <- unstd.posterior / sum(unstd.posterior)
  
  plot(p_grid, posterior, type = "b", xlab = "probability of water",
       ylab = "posterior probability")
}
grid_post_step_prior(w = 3, n = 3)
grid_post_step_prior(w = 3, n = 4)
grid_post_step_prior(w = 5, n = 7)
```

2M3. Suppose there are two globes: Mars and Earth. Earth glove is 70% covered in water. Mars is 100% land. One globe was tossed and produced "L." Show that the posterior probability that the globe was Earth, conditional on seeing land (Pr(Earth\|land)) = 0.23

```{r}

# Answer: 
# Pr(Earth | land) = Pr(land | Earth) * Pr(Earth) / Pr(land)
# Pr(land) = all the ways we could get land 
# Pr(land) = Pr(land|Earth)*Pr(Earth) + Pr(land|mars)*Pr(Mars)

0.3 * 0.5 / ((0.3*0.5) + (1 * 0.5))

```

2M4: Suppose you have a deck with 3 cards. Each card has two sides, each side is black or white. One card is BB, one card is BW, the third is WW. A card has one black side. What is the probability that the other side is black?

```{r}
# answer

# Total possibilities
# B - B
# B - W
# W - W 

# if we observe a B face up we know we have either 
# B - B
# B - W

# We know that one of the four faces we've seen is a B. 2/3 of the unknown faces are Bs, so 2/3 chance it's a BB


```

2M5. Suppose there are now four cards: BB, BW, WW, and BB. Calculate the probability of B\|B

```{r}
# Answer 
# If we observe a B we know we have one of: 
# B-B
# B-B
# B-W

# Now 4/5 unknown faces are B, so probability is 4/5 

```

2M6. Imagine that black ink is heavy, so its less likely to grab a black sided card. Assume 3 cards, and for every way to pull a B/B, there are 2 ways to pull a B/W and three ways to pull W/W

```{r}
# Observed Data = B 

# Ways to observe it (B/W is twice as likely): 
# B/B
# B/W
# B/W 

# Now there are 2 ways to get B from a BW card and 2 ways to get B from a BB card. So 0.5
```

2M7. Assume the card problem, first card is B, second card is W. What is the probability that the first card is BB?

```{r}
#Answer: There are two hypotheses

# h1 = BB is first 
# h2 = BW is first 
# 
# Ways to produce h1:
# BB (2 ways to get b) * [BW (1) + WW (2)] = 6

# Ways to produce h2: 
# BW(1 way to get a b) * WW (2) = 2 

# total ways = 8
# BB ways = 6/8 

```

2H1 Suppose there are two species of panda bear. Both are equally common in the wild and live in the same places. They look exactly alike and eat the same food, and there is yet no genetic assay capable of telling them apart. They differ however in their family sizes. Species A gives birth to twins 10% of the time, otherwise birthing a single infant. Species B births twins 20% of the time, otherwise birthing singleton infants. Assume these numbers are known with certainty, from many years of field research.

Now suppose you are managing a captive panda breeding program. You have a new female panda of unknown species, and she has just given birth to twins. What is the probability that her next birth will also be twins?

```{r}
# Answer: What is the Pr of (twins2 | twins1)

# Pr(t2|t1) =  Pr(t1 and t2) / Pr(t1)

# Pr(t1) = 0.5 * .1 + 0.5 * .2 =  0.15
# Pr(t1 and t2) =  .5*(.1*.1) +  .5*(.2*.2) = 0.025
.025/.15
```

2H2. Recall all the facts from the problem above. Now compute the probability that the panda we have is from species A, assuming we have observed only the first birth and that it was twins.

```{r}
# Answer
# Pr(A|twins) = Pr(twins|A)*Pr(A) / Pr(twins)
Pr_twins = .1*.5 + .2*.5

.5 * .1 / .15
```

2H3. Continuing on from the previous problem,suppose the same panda mother has a second birth and that it is not twins, but a singleton infant. Compute the posterior probability that this panda is species A.

```{r}

# Answer: we use the previous data (1/3) as the Pr(A), aka our prior 

prior = 1/3
#Pr(A | singleton) = Pr(singleton|A) * prior / Pr(singleton)
.9 * prior / .85

```

2H4. A common boast of Bayesian statisticians is that Bayesian inference makes it easy to use all of the data, even if the data are of different types.

So suppose now that a veterinarian comes along who has a new genetic test that she claims can identify the species of our mother panda. But the test, like all tests, is imperfect. This is the informa- tion you have about the test: \* The probability that it correctly identifies species A panda is 0.8 \* The probability that it correctly identifies species B panda is 0.65

The vet administers the test to your panda and tells you that the test is positive for species A. First ignore your previous information from the births and compute the posterior probability that your panda is species A. Then redo your calculation, now using the birth data as well.

```{r}
# Answer, using just the genetic data 
# Pr(posA | A) = 0.8
# Pr(posA | B) = 0.35

# First, find Pr(A | pos) = Pr(pos|A) * Pr(A) / Pr(pos)

(.8*.5)/(.8*.5 + .35*.5)

```

```{r}
#Then, use the birth data as well. 

# Now, we'll update Pr(A) to reflect birth data telling us that it's 
# not likely an A (p = 0.352)
(.8*.352)/(.8*.352 + .35*.65)


```
