---
title: "Chapter_4.7_Practice_exercises"
format: html
editor: visual
embed-resources: true
---

### Chapter 4.7, end of chapter exercises. 

```{r, message= F, warning = F}
library(dplyr)
library(rethinking)
```

4E1: In the model definition below, which line is the liklihood?  
Answer: yi ~ Normal(mu, sigma)

4E2: In the model definition just above, how many parameters 
are in the posterior?  
Answer: 2 

4E3: Using the model definition above, write down the appropriate form of Bayes theorem that includes the proper likelihood and priors.  
Answer: see https://xcelab.net/rmpubs/rethinking/rethinking_solutions_2.pdf but the start of it is rephrasing the model as 
Pr(sigma and u |y)

4E4: What line is the linear model?  
mu = alpha + beta*xi

4E5: How many parameters are in the posterior?  
three.

4M1: For the model definition  below, simulate observed y 
values from the prior (not the posterior) 
SMM note: In his answers RM uses a uniform distribution for  the sigma... why? 
```{r}
# model 
# yi ~ Normal(mu, sigma)
# mu ~ Normal(0, 10)
# sigma ~ Exponential (1) 
mus <- rnorm(10000, mean = 0, sd = 10)
sigmas <- runif(n = 10000, rate =1)

ys <- rnorm(10000, mean = mus, sd = sigmas)
hist(ys)
```
4M2. Translate the model above into a quap formula
```{r}
m4M2 <- 
    alist(
        y ~ dnorm( mu , sigma ) ,
        mu ~ dnorm( 0 ,10 ) ,
        sigma ~ dunif( 0 , 10 )
    )

```

4M3. Translate the quap model formula into a math definition 
SMM note: I don't understand the translations between math and R distributions 
```{r}
# y ~ Normal(mu, sigma), 
# mu <- alpha + beta*xi
# alpha ~ Normal(0, 10)
# beta ~ Exponential (0,1)
# sigma ~  Uniform (1)  
```

4M4. A sample of students is measured for height for 3 years. After the 3rd year you want to fit a regression predicting height by year. Write the math model definition. 

```{r}
# heighti ~ normal(mu, sigma)
# mui = alpha + beta * yeari
# alpha ~ normal (100, 10) ;  say that mean in year 1 is 100, sd 10
# beta ~ normal (0, 10) 
# sigma ~ uniform(0, 50)
```

4M5. Every student gets taller each year- how should we change our prior? 
```{r}
#beta = lognormal(0, 1)
```
4M6.  suppose I tell you that the variance for students of the same age never gets more than 64 cm. 

```{r}

# Sigma ~ unif (0, 64)
```

4H1. The weights listed below were recordded but heights were not. Provide heights and 89% intervals for each individual

```{r}
# get the data 
data(Howell1)
dat <- Howell1 %>% filter(age >= 18)
head(dat)
xbar <- mean(dat$weight)
# Make the model 
mod_height <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b*(weight),
    a ~ dnorm(100, 100),
    b ~ dlnorm(0, 10),
    sigma ~ dunif(0, 50)
  ), data = dat
  )

# use link to predict values 
samples <- link(mod_height, data = data.frame(weight = c(46.95, 43.72, 64.78, 32.59, 54.63)), n = 1e5)

apply(samples, 2, mean) # mean predicted values 
apply(samples, 2, function (x)  HPDI(x, prob = 0.9))

# HPDI didn't work the same as in the book. try something else? 

post<- extract.samples(mod_height)
head(post)
y <- rnorm( 1e5 , post$a + post$b*46.95 , post$sigma )
HPDI(y, prob = 0.9) # this gives the same answer as book
range(y)
range(samples[,1])

```
Ok I figured out the difference. Link pulls samples from the posterior for the three parameters that we estimated (a, b, and sigma). It then uses a and b in our linear model to predict height, assuming that our best guess of height is = a + b* weight. 

The book answer incorporates sigma, and says that height is best understood as normally distributed with a mean of the linear regression, but with sd = sigma. So there's more variation in it. Confusing. 

4H2. Model people under 18. 

```{r}

data(Howell1)
kids <- Howell1 %>% filter(age < 18)
head(kids)
xbar <- mean(kids$weight)
# find some reasonable priors
mean(kids$weight)

# Make the model, make the priors work 
mod_height2 <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b*(weight),
    a ~ dnorm(18, 18),
    b ~ dlnorm(0, 4),
    sigma ~ dunif(0, 50)
  ), data = kids
  )

# interpret model
precis(mod_height2)

```

A). For every 10 increase in weight, height will increase by ~27 cm. 

```{r}
#B) Plot the raw data, height on the vertical, weight on # the horizonal

# Get the regression line 
link_mod <- link(mod_height2)
head(link_mod)
samps <- extract.samples(mod_height2)
# extract the mean slope and mean intercept
mean_params <- apply(samps, 2, mean)

# extract the .89 interval for the mean 
mu.PI <- apply(samps, 2, PI, prob = .89)

# extract the 89% interval for predicted heights 

plot(height ~ weight, data = kids, col = "lightblue")
mapply(function(a,b) abline(a = a, b= b), samps$a[1:10], samps$b[1:10]) # Plot a bunch of lines from the posterior

abline(a = mean_params[1], b = mean_params[2], col = "red") # add a red line indicating the mean regression

# Get the 89% interval for the mean 


w.seq <- seq(from=1,to=45,length.out=50)
mu <- sapply( w.seq , function(z) mean( samps$a + samps$b*z ) )
mu.ci <- sapply( w.seq , function(z)
HPDI( samps$a + samps$b*z , prob=0.9 ) )
pred.ci <- sapply( w.seq , function(z)
HPDI( rnorm(10000,samps$a + samps$b*z,samps$sigma) , 0.9 ) )




```

